### 2020-12-28 latex/acrobot.tex
Fixed some typos caught by mother

Now consider the solution \(x(t,\eta_0,I)\) to ~\eqref{eqn:khalil-setup}.
Performing a first-order Taylor series expansion at \(I = 0\), we get
\(x(t,\eta_0,I) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\),
where the remainder term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie, 
\[
    \lim \limits_{I \to 0} \frac{R(t,\eta_0,I)}{I} = 0
    .
\]
One can recover \(x_1(t,\eta_0)\) as the solution to the scalar
time-varying ODE
\begin{equation}\label{eqn:khalil-perturbation-firstorder}
    \begin{cases}
        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t,\eta_0),0)x_1 + \pdiff{f}{I}(t,x_0(t,\eta_0),0)
        , \\
        x_1(t_0,\eta_0) = 0
        .
    \end{cases}
\end{equation}

@@

Since \(\dot{\alpha} > 0\), there is a well-defined Poincar\'{e} map 
describing how the pseudo-radius \(\mu\) changes each time the orbit of \((q_u(t),p_u(t))\)
intersects the \(q_u\)-axis, \ie, when \(\alpha = 0\) or \(\alpha = \pi\).
Using our time-scaled system ~\eqref{eqn:muhat-dot}, we can define this
Poincar\'{e} map as the change in \(\hat{\mu}\) when \(\alpha\) increases by
\(\pi\):
\begin{align*}
    P_\mathcal{O} : \,]0,\pi[ &\rightarrow \R_+,
    \\
    \mu_0 &\mapsto \hat{\mu}(\pi,\mu_0,I)
    .
\end{align*}
The Poincar\'{e} map expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi a(\sigma,\mu_0)d\sigma
    + R(\pi,\mu_0,I)
    .
\]
Let us define
\begin{equation}\label{eqn:acrobot-Qmu}
    Q(\mu_0) := \int \limits_0^\pi a(\sigma,\mu_0)d\sigma
    .
\end{equation}
Note that \(K\) is a positive constant which contains the acrobot's physical
parameters \(m\), \(g\), \(l\), and \(\bar{q}_a\), 
while \(a(\alpha,\mu)\) is adimensional. 
This means \(Q(\mu_0)\) is identical for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive for any acrobot.

@@

Suppose we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\).
One full rotation amounts to \(\beta\) traversing \(2\pi\) rad in a clockwise
direction; that is, when \(\beta\) goes from \(0\) to \(\sign{\rho_0}2\pi\).
Since \(\dot{\beta}\) does not change sign on \(\mathcal{R}\),
there is a well-defined Poincar\'{e} map describing how the pseudo-radius
\(\rho\) changes each time the orbit \((q(t),p(t))\) hits the \(p_u\) axis, \ie,
every time \(\beta\) changes by \(2\pi\).
We define this Poincar\'{e} map as
\[
    P_\mathcal{R}(\rho_0) := \hat{\rho}\left(\sign{\rho_0}2\pi,\rho_0,I\right)
    ,
\]
which expands into
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
    + R(\sign{\rho_0}2\pi,\rho_0,I)
    .
\]

Let us pause here for a moment to remember what we are trying to accomplish.
The second part of Theorem \ref{thm:acrobot-energy-stabilization} states that,
under suitable conditions on the integral of \(b(\beta,\rho_0)\), 
the acrobot will gain energy on 
\[
    \mathcal{O}_2(\bar{\rho}) := \left\{ (q_u,p_u) \in \SxR
    \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    ,
\]
for some \(\bar{\rho} > \sqrt{60m^2gl^3}\).
Recall from Section \ref{sec:acrobot-proof-o1} that any acrobot initialized in
\(\mathcal{O}_1\) will eventually remain in the rotation region \(\mathcal{R}\).
What remains, then, is to prove that the acrobot's orbits will also gain
energy on
\[
    \mathcal{R}_{\bar{\rho}} := \mathcal{R} \cap \mathcal{O}_2(\bar{\rho})
    .
\]
This can be proven by analyzing the Poincar\'{e} map \(P_\mathcal{R}(\rho_0)\).
Looking at Figure \ref{fig:acrobot-rhobar-regions}, it is clear that
\(P_\mathcal{R}\) has the domain
\[
    D := \left[-\bar{\rho},-\sqrt{60m^2gl^3}\right[ \, 
    \cup \, 
    \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
    ,
\]
which is the intersection of \(\mathcal{R}_{\bar{\rho}}\) with the \(p_u\)-axis.
We label the part of \(D\) contained in the negative \(p_u\)-axis as
\[
    D^- :=\left[-\bar{\rho},-\sqrt{60m^2gl^3}\right[
    ,
\]
and the part contained in the positive \(p_u\)-axis as
\[
    D^+ := \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
    ,
\]
so that \(D = D^- \cup D^+\).
If \(P_\mathcal{R}(\rho_0)\) is always further from the origin than 
\(\rho_0\) (\ie~ it is ``expanding" on \(D\)), the acrobot will be gaining energy
on \(\mathcal{O}_2(\bar{\rho})\).

@@

Here is a summary of the proof:
\begin{enumerate}
    \item We found pseudo-polar coordinates \((\beta,\rho)\) adapted to level
        sets of the nominal pendulum's mechanical energy on \(\mathcal{R}\),
        where \(\beta \in \Sone\) is a pseudo-angle and
        \(\rho \in D\) is a pseudo-radius.
    \item We showed there exists a value \(I_1\) small enough where
        \(\dot{\beta} > 0\) everywhere on \(\mathcal{R}\).
    \item Using \(\beta\) as a time variable, we found the evolution of the
        time-scaled pseudo-radius \(\hat{\rho}\) and expanded it into
        ~\eqref{eqn:acrobot-rhohat-approx} using perturbation theory.
    \item Taking this expanded solution, we defined and expanded the
        Poincar\'{e} map \(P_\mathcal{R}\) on the domain 
        \(D = D^- \cup D^+\) which shows how the acrobot's orbits behave on
        \(\mathcal{R}_{\bar{\rho}}\).
    \item Using the Poincar\'{e} map to analyze the discrete-time system
        ~\eqref{eqn:rhohat-discrete}, we found values \(I_2,I_4\) bounding the
        remainder term \(R(\beta,\rho_0,I)\) on \(D^+\) and \(D^-\)
        (respectively), then found values \(I_3, I_5\) which drive all orbits of
        ~\eqref{eqn:rhohat-discrete} towards \(\abs{\rho} = \bar{\rho}\).
\end{enumerate}
Recall from Section \ref{sec:acrobot-proof-o1} that there exists \(I^\star > 0\)
making orbits escape \(\bar{\mathcal{O}}_1\) into 
\(\mathcal{R}\). 
Choosing \(I \in \,]0,\min\{I_3,I_5,I^\star\}]\)
guarantees energy injection on \(\mathcal{O}_2(\bar{\rho})\).
Likewise, choosing \(I \in [-\min\{I_3,I_5,I^\star\},0[\) guarantees energy
dissipation on \(\mathcal{O}_2(\bar{\rho})\).

This analysis, together with that of section \ref{sec:acrobot-proof-o1},
completes the proof of Theorem \ref{thm:acrobot-energy-stabilization}.

>>>>

Now consider the solution \(x(t,\eta_0,I)\) to ~\eqref{eqn:khalil-setup}.
Performing a first-order Taylor series expansion at \(I = 0\), we get
\(x(t,\eta_0,I) = x_0(t,\eta_0) + I x_1(t,\eta_0) + R(t,\eta_0,I)\),
where the remainder term \(R(t,\eta_0,I)\) is smooth and \(O(I^2)\), \ie 
\[
    \lim \limits_{I \to 0} \frac{R(t,\eta_0,I)}{I} = 0
    .
\]
One can recover \(x_1(t,\eta_0)\) as the solution to the scalar
time-varying ODE
\begin{equation}\label{eqn:khalil-perturbation-firstorder}
    \begin{cases}
        \dot{x}_1 = \pdiff{f}{x}(t,x_0(t,\eta_0),0)x_1 + \pdiff{f}{I}(t,x_0(t,\eta_0),0)
        , \\
        x_1(t_0,\eta_0) = 0
        .
    \end{cases}
\end{equation}

@@

Since \(\dot{\alpha} > 0\), there is a well-defined Poincar\'{e} map 
describing how the pseudo-radius \(\mu\) changes each time the orbit of \((q_u(t),p_u(t))\)
intersects the \(q_u\)-axis, \ie when \(\alpha = 0\) or \(\alpha = \pi\).
Using our time-scaled system ~\eqref{eqn:muhat-dot}, we can define this
Poincar\'{e} map as the change in \(\hat{\mu}\) when \(\alpha\) increases by
\(\pi\):
\begin{align*}
    P_\mathcal{O} : \,]0,\pi[ &\rightarrow \R_+,
    \\
    \mu_0 &\mapsto \hat{\mu}(\pi,\mu_0,I)
    .
\end{align*}
The Poincar\'{e} map expands into
\[
    P_\mathcal{O}(\mu_0) = \mu_0 + I K \int\limits_0^\pi a(\sigma,\mu_0)d\sigma
    + R(\pi,\mu_0,I)
    .
\]
Let us define
\begin{equation}\label{eqn:acrobot-Qmu}
    Q(\mu_0) := \int \limits_0^\pi a(\sigma,\mu_0)d\sigma
    .
\end{equation}
Note that \(K\) is a positive constant which contains the acrobot's physical
parameters \(m\), \(g\), \(l\), and \(\bar{q}_a\), 
while \(a(\alpha,\mu)\) is adimensional. 
This means \(Q(\mu_0)\) is identical for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive for any acrobot.

@@

Suppose we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\).
One full rotation amounts to \(\beta\) traversing \(2\pi\) rad in a clockwise
direction; that is, when \(\beta\) goes from \(0\) to \(\sign{\rho_0}2\pi\).
Since \(\dot{\beta}\) does not change sign on \(\mathcal{R}\),
there is a well-defined Poincar\'{e} map describing how the pseudo-radius
\(\rho\) changes each time the orbit \((q(t),p(t))\) hits the \(p_u\) axis, \ie
every time \(\beta\) changes by \(2\pi\).
We define this Poincar\'{e} map as
\[
    P_\mathcal{R}(\rho_0) := \hat{\rho}\left(\sign{\rho_0}2\pi,\rho_0,I\right)
    ,
\]
which expands into
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
    + R(\sign{\rho_0}2\pi,\rho_0,I)
    .
\]

Let us pause here for a moment to remember what we are trying to accomplish.
The second part of Theorem \ref{thm:acrobot-energy-stabilization} states that,
under suitable conditions on the integral of \(b(\beta,\rho_0)\), 
the acrobot will gain energy on 
\[
    \mathcal{O}_2(\bar{\rho}) := \left\{ (q_u,p_u) \in \SxR
    \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    ,
\]
for some \(\bar{\rho} > \sqrt{60m^2gl^3}\).
Recall from Chapter \ref{sec:acrobot-proof-o1} that any acrobot initialized in
\(\mathcal{O}_1\) will eventually remain in the rotation region \(\mathcal{R}\).
What remains, then, is to prove that the acrobot's orbits will also gain
energy on
\[
    \mathcal{R}_{\bar{\rho}} := \mathcal{R} \cap \mathcal{O}_2(\bar{\rho})
    .
\]
This can be proven by analyzing the Poincar\'{e} map \(P_\mathcal{R}(\rho_0)\).
Looking at Figure \ref{fig:acrobot-rhobar-regions}, it is clear that
\(P_\mathcal{R}\) has the domain
\[
    D := \left[-\bar{\rho},-\sqrt{60m^2gl^3}\right[ \, 
    \cup \, 
    \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
    ,
\]
which is the intersection of \(\mathcal{R}_{\bar{\rho}}\) with the \(p_u\)-axis.
We label the part of \(D\) contained in the negative \(p_u\)-axis as
\[
    D^- :=\left[-\bar{\rho},-\sqrt{60m^2gl^3}\right[
    ,
\]
and the part contained in the positive \(p_u\)-axis as
\[
    D^+ := \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
    ,
\]
so that \(D = D^- \cup D^+\).
If \(P_\mathcal{R}(\rho_0)\) is always further from the origin than 
\(\rho_0\) (\ie it is ``expanding" on \(D\)), the acrobot will be gaining energy
on \(\mathcal{O}_2(\bar{\rho})\).

@@

Here is a summary of the proof:
\begin{enumerate}
    \item We found pseudo-polar coordinates \((\beta,\rho)\) adapted to level
        sets of the nominal pendulum's mechanical energy on \(\mathcal{R}\),
        where \(\beta \in \Sone\) is a pseudo-angle and
        \(\rho \in D\) is a pseudo-radius.
    \item We showed there exists a value \(I_1\) small enough where
        \(\dot{\beta} > 0\) everywhere on \(\mathcal{R}\).
    \item Using \(\beta\) as a time variable, we found the evolution of the
        time-scaled pseudo-radius \(\hat{\rho}\) and expanded it into
        ~\eqref{eqn:acrobot-rhohat-approx} using perturbation theory.
    \item Taking this expanded solution, we defined and expanded the
        Poincar\'{e} map \(P_\mathcal{R}\) on the domain 
        \(D = D^- \cup D^+\) which shows how the acrobot's orbits behave on
        \(\mathcal{R}_{\bar{\rho}}\).
    \item Using the Poincar\'{e} map to analyze the discrete-time system
        ~\eqref{eqn:rhohat-discrete}, we found values \(I_2,I_4\) bounding the
        remainder term \(R(\beta,\rho_0,I)\) on \(D^+\) and \(D^-\)
        (respectively), then found values \(I_3, I_5\) which drive all orbits of
        ~\eqref{eqn:rhohat-discrete} towards \(\abs{\rho} = \bar{\rho}\).
\end{enumerate}
Recall from Chapter \ref{sec:acrobot-proof-o1} that there exists \(I^\star > 0\)
making orbits escape \(\bar{\mathcal{O}}_1\) into 
\(\mathcal{R}\). 
Choosing \(I \in \,]0,\min\{I_3,I_5,I^\star\}]\)
guarantees energy injection on \(\mathcal{O}_2(\bar{\rho})\).
Likewise, choosing \(I \in [-\min\{I_3,I_5,I^\star\},0[\) guarantees energy
dissipation on \(\mathcal{O}_2(\bar{\rho})\).

This analysis, together with that of Chapter \ref{sec:acrobot-proof-o1},
completes the proof of Theorem \ref{thm:acrobot-energy-stabilization}.

### 2020-12-28 latex/vlp.tex
Fixed some typos caught by mother

We now prove that almost all solutions of the reduced
dynamics~\eqref{eqn:vlp-hamiltonian} escape compact sets in finite
time, \ie, that for each compact subset \(K\) of 
\(\SxR\), and for almost every condition 
\( (q_0,p_0) \in K \), there exists \( T > 0\) such that
\( (q(t),p(t)) \notin K \) for all \( t > T \), where \( (q(t),p(t)) \)
denotes the solution of~\eqref{eqn:vlp-hamiltonian} with 
\( (q(0),p(0)) = (q_0,p_0)\). 
In what follows, we denote by \( x = (q,p) \) the state of the reduced
dynamics~\eqref{eqn:vlp-hamiltonian}. 
Also, we denote by \(\Pi^+\) the stable manifold of the equilibrium \((\pi,0)\)
for system~\eqref{eqn:vlp-hamiltonian}, defined as
\[
   \Pi^+ := \left\{ x(0) \in \SxR \mid
   \lim_{t \to \infty} x(t) = (\pi, 0) \right\}.
\]
The linearization of system~\eqref{eqn:vlp-hamiltonian} at the
equilibrium \((\pi,0)\) has a system matrix given by
\begin{equation}\label{eqn:linearized-vlp}
   \begin{bmatrix}
      0 & -\frac{1}{mL(0)^2} \\
      -m g L(0) & 0 \\
   \end{bmatrix}
   ,
\end{equation}
whose spectrum is \(\{\pm \sqrt{{g}/{L(0)}}\}\).  
Since one eigenvalue of ~\eqref{eqn:linearized-vlp} is positive and one is
negative, the stable manifold theorem implies that the stable manifold 
\( \Pi^+ \) is a one-dimensional immersed submanifold of
\(\SxR\), and thus is a set of measure zero
\cite{stable_manifold}.

@@

The average weight of a 25 year old Estonian man is \(80.17\)kg
\cite{estonian-body-data}, so we set the VLP mass \(m\) to this value.
To find our upper and lower bounds on length, we first need to choose the size
of a swing.
The common model of a Kiiking swing is between \(3-4\)m tall
\cite{kiiking-sale}, so we take the length to be \(3\)m.
The average 25 year old Estonian is \(1.8\)m tall \cite{estonian-body-data}, and
we assume their squat height is half that at \(0.9\)m tall.
We further assume that their center of mass (COM) is always located at half
their current height, which means their standing COM is \(0.9\)m from the ground
while their squatting COM is \(0.45\)m from the ground.
This means the distance from their standing COM to the pivot is \(2.1\)m, while
the distance from their squatting COM to the pivot is \(2.55\)m.
Hence, we set \(\underbar{l} = 2.1\) and \(\overbar{l} = 2.55\).
We initialize a VLP with these parameters at 
\((q,p) = \left(\frac{\pi}{32},0\right)\) and plot the resulting orbits
in Figure \ref{fig:vlp-sim-orbit}.

\begin{figure}
   \centering
   \includegraphics[width=0.8\textwidth]{images/kiiking_orbits.png}
   \caption{The orbits of the time-optimal controller \(L^\star(\theta)\)
      (black), the differentiable controller \(L_T(\theta)\) (magenta),
      and the smooth VNHC \(L_\frac{\pi}{2}(\theta)\) (cyan) employed by a 25
      year old Estonian male on a 3m tall Kiiking swing. All the controllers are
      initialized at \(q(0) = \frac{\pi}{32}\), and they each steadily gain
      energy until they reach a rotation.}
   \label{fig:vlp-sim-orbit}
\end{figure}

@@

\begin{figure}
   \centering
   \includegraphics[width=0.9\textwidth]{images/E_kiiking_ltheta.png}
   \caption{The energy \(E_\text{avg}(q,p)\) of a 25 year old Estonian male on a
      3m tall Kiiking swing who employs the smooth VNHC
   \(L_\frac{\pi}{2}(\theta)\).}
   \label{fig:vlp-sim-energy}
\end{figure}

>>>>

We now prove that almost all solutions of the reduced
dynamics~\eqref{eqn:vlp-hamiltonian} escape compact sets in finite
time, \ie that for each compact subset \(K\) of 
\(\SxR\), and for almost every condition 
\( (q_0,p_0) \in K \), there exists \( T > 0\) such that
\( (q(t),p(t)) \notin K \) for all \( t > T \), where \( (q(t),p(t)) \)
denotes the solution of~\eqref{eqn:vlp-hamiltonian} with 
\( (q(0),p(0)) = (q_0,p_0)\). 
In what follows, we denote by \( x = (q,p) \) the state of the reduced
dynamics~\eqref{eqn:vlp-hamiltonian}. 
Also, we denote by \(\Pi^+\) the stable manifold of the equilibrium \((\pi,0)\)
for system~\eqref{eqn:vlp-hamiltonian}, defined as
\[
   \Pi^+ := \left\{ x(0) \in \SxR \mid
   \lim_{t \to \infty} x(t) = (\pi, 0) \right\}.
\]
The linearization of system~\eqref{eqn:vlp-hamiltonian} at the
equilibrium \((\pi,0)\) has a system matrix given by
\begin{equation}\label{eqn:linearized-vlp}
   \begin{bmatrix}
      0 & -\frac{1}{mL(0)^2} \\
      -m g L(0) & 0 \\
   \end{bmatrix}
   ,
\end{equation}
whose spectrum is \(\{\pm \sqrt{{g}/{L(0)}}\}\).  
Since one eigenvalue of ~\eqref{eqn:linearized-vlp} is positive and one is
negative, the stable manifold theorem implies that the stable manifold 
\( \Pi^+ \) is a one-dimensional immersed submanifold of
\(\SxR\), and thus is a set of measure zero
\cite{stable_manifold}.

@@

The average weight of a 25-year-old Estonian man is \(80.17\)kg
\cite{estonian-body-data}, so we set the VLP mass \(m\) to this value.
To find our upper and lower bounds on length, we first need to choose the size
of a swing.
The common model of a Kiiking swing is between \(3-4\)m tall
\cite{kiiking-sale}, so we take the length to be \(3\)m.
The average 25-year-old Estonian is \(1.8\)m tall \cite{estonian-body-data}, and
we assume their squat height is half that at \(0.9\)m tall.
We further assume that their center of mass (COM) is always located at half
their current height, which means their standing COM is \(0.9\)m from the ground
while their squatting COM is \(0.45\)m from the ground.
This means the distance from their standing COM to the pivot is \(2.1\)m, while
the distance from their squatting COM to the pivot is \(2.55\)m.
Hence, we set \(\underbar{l} = 2.1\) and \(\overbar{l} = 2.55\).
We initialize a VLP with these parameters at 
\((q,p) = \left(\frac{\pi}{32},0\right)\) and plot the resulting orbits
in Figure \ref{fig:vlp-sim-orbit}.

\begin{figure}
   \centering
   \includegraphics[width=0.8\textwidth]{images/kiiking_orbits.png}
   \caption{The orbits of the time-optimal controller \(L^\star(\theta)\)
      (black), the differentiable controller \(L_T(\theta)\) (magenta),
      and the smooth VNHC \(L_\frac{\pi}{2}(\theta)\) (cyan) employed by a
      25-year-old Estonian male on a 3m tall Kiiking swing. All the controllers
      are initialized at \(q(0) = \frac{\pi}{32}\), and they each steadily gain
      energy until they reach a rotation.}
   \label{fig:vlp-sim-orbit}
\end{figure}

@@

\begin{figure}
   \centering
   \includegraphics[width=0.9\textwidth]{images/E_kiiking_ltheta.png}
   \caption{The energy \(E_\text{avg}(q,p)\) of a 25-year-old Estonian male on a
      3m tall Kiiking swing who employs the smooth VNHC
   \(L_\frac{\pi}{2}(\theta)\).}
   \label{fig:vlp-sim-energy}
\end{figure}
